---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Import the RAM Legacy Stock Assessment Database into R

[![Travis-CI Build Status](https://travis-ci.org/seananderson/ramlegacy.svg?branch=master)](https://travis-ci.org/seananderson/ramlegacy)

This package does one thing: it downloads a Microsoft Access copy of the [RAM Legacy Stock Assessment Database](http://ramlegacy.org) and converts it to a local sqlite3 database. This makes it easy to [work with dplyr/dbplyr](https://cran.r-project.org/web/packages/dbplyr/vignettes/dbplyr.html), for example. The `make_ramlegacy()` function also leaves a copy of `.csv` files for each table in the database in the R working directory if you'd prefer to work with those.

**Note that you must have the utility `mdb-tables` installed** and in your path from `mdbtools`. This utility provides tools for extracting Access databases. You can find installation instructions at <http://mdbtools.sourceforge.net>. If you're on OS X and using homebrew, you can install it with `brew install mdbtools`.

### Example use

```{r knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE)
```

Install the package:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("seananderson/ramlegacy")
```

Cache and convert the database:

```{r}
library("ramlegacy")
make_ramlegacy()
```

Work with the data:

```{r}
library("dplyr")
ram <- src_sqlite("ramlegacy.sqlite3")
ram # see the available tables
```

Access the `area` table:

```{r}
tbl(ram, "area")
```

Join the time series `ts` and `stock` tables on the `stockid` column:

```{r}
ts <- tbl(ram, "timeseries")
stock <- tbl(ram, "stock")
select(stock, stockid, scientificname, commonname, region) %>%
  inner_join(ts)
```

Note that because you are working with dplyr and a database, you will need to use `dplyr::collect()` once you want to load the data into a local data frame. For example:

```{r}
tbl(ram, "area") %>% 
  collect()
```

For safety, you may want to use `dplyr::collect(n = Inf)` to return all rows of data, not just the minimum default number. In this case it won't make a difference.

```{r}
tbl(ram, "area") %>% 
  collect(n = Inf)
```
