# Download and convert the RAM Legacy Stock Assessment Database in R

This package does one thing right now: it downloads the latest Microsoft Access version of the [RAM Legacy Stock Assessment Database](http://ramlegacy.org) (version 2.5) and converts it to a local sqlite3 database. This makes it easy to [work with dplyr](http://cran.r-project.org/web/packages/dplyr/vignettes/databases.html), for example. The `make_ramlegacy()` function also leaves a copy of `.csv` files for each table in the database in the R working directory.

Perhaps eventually I'll add additional functionality, but for now I just wanted a reproducible way to bring the latest copy of the database into a local (non-Access) database that I could use with R.

### Example use

```{r knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE)
```

Install the package:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("seananderson/ramlegacy")
```

Cache and convert the database:

```{r}
library("ramlegacy")
make_ramlegacy()
```

Work with the data:

```{r}
library("dplyr")
ram <- src_sqlite("ramlegacy.sqlite3")
ram
```

```{r}
area <- tbl(ram, "area")
glimpse(area)
ts <- tbl(ram, "timeseries")
stock <- tbl(ram, "stock")
select(stock, stockid, scientificname, commonname, region) %>%
  inner_join(select(ts, -tsid, -assessid, -stocklong)) %>%
  glimpse
```